---
title: "DESeq2"
output: pdf_document
date: "2024-01-25"
---

```{r}
setwd("/Users/tiknokeziah/Desktop/Keziah/School/Research Project/Data/Count Data")

library(tidyr)
library(dplyr)
library(limma)
library(edgeR)
library(DESeq2)
library(reshape2)
library(ggplot2)
library(rstatix)
library(ggpubr)
```

# Upload RData
```{r}
gene_list <- readRDS("gene_list.RData")
all_counts <- readRDS("All_Counts.RData")

metadata <- readRDS('metadata.RData')
metadata$sample_name <- rownames(metadata)

# Normalized counts
dds <- readRDS("dds.normal_vs_ko.RData")
norm_counts <- data.frame(counts(dds, normalized=TRUE))
```

# Data manipulation for ggplot DMSO vs DMSO_RB1-KD
```{r}
df.res <- readRDS("df.res.RData")
df.res.fus <- readRDS("df.res.fus.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot DMSO vs Abema
```{r}
df.res <- readRDS("df.res.abema.RData")
df.res.fus <- readRDS("df.res.fus_abema.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot DMSO vs Abema_RB1-KD
```{r}
df.res <- readRDS("df.res.abema.ko.RData")
df.res.fus <- readRDS("df.res.fus_abema.ko.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot Abema vs DMSO_RB1-KD
```{r}
df.res <- readRDS("df.res.abema_dmso.ko.RData")
df.res.fus <- readRDS("df.res.fus_abema_dmso.ko.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot Abema vs Abema_RB1-KD
```{r}
df.res <- readRDS("df.res.abema_abema.ko.RData")
df.res.fus <- readRDS("df.res.fus_abema_abema.ko.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot DMSO_RB1-KD vs Abema_RB1-KD
```{r}
df.res <- readRDS("df.res.dmso.ko_abema.ko.RData")
df.res.fus <- readRDS("df.res.fus_dmso.ko_abema.ko.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot DMSO vs Abema
```{r}
df.res <- readRDS("df.res.dmso_vs_abema.RData")
df.res.fus <- readRDS("df.res.fus.dmso_vs_abema.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$treatment <- factor(melt_norm$treatment, levels = c('DMSO','abemaciclib'))
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$treatment <- factor(melt_norm_fus$treatment, levels = c('DMSO','abemaciclib'))
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# Data manipulation for ggplot Normal vs KD
```{r}
df.res <- readRDS("df.res.normal_vs_ko.RData")
df.res.fus <- readRDS("df.res.fus.normal_vs_ko.RData")
df.res.fus$hgnc_symbol <- rownames(df.res.fus)

# Subsetting normalized counts that have an ordered significant pvalues
norm_counts.sig <- norm_counts[na.omit(head(df.res[order(df.res$padj),], 4))$entrezgene_id,]

norm_counts.sig.fus <- norm_counts[na.omit(head(df.res.fus[order(df.res.fus$padj),], 4))$hgnc_symbol,]

# Manipulating the data
melt_norm <- melt(as.matrix(norm_counts.sig))
colnames(melt_norm) <- c("entrezgene_id", "sample_name", "norm_counts")

melt_norm_fus <- melt(as.matrix(norm_counts.sig.fus))
colnames(melt_norm_fus) <- c("hgnc_symbol", "sample_name", "norm_counts")

# Merging the metadata and gene list to the melt_norm to add new_treatment and symbol columns
melt_norm <- merge(melt_norm, metadata, by="sample_name",all.x=T)
melt_norm <- merge(melt_norm, gene_list, by="entrezgene_id", all.x=T)
melt_norm$norm_counts <- log2(melt_norm$norm_counts)
melt_norm$group <- factor(melt_norm$group, levels = c('Normal','KD'))
melt_norm$new_treatment <- factor(melt_norm$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

melt_norm_fus <- merge(melt_norm_fus, metadata, by="sample_name",all.x=T)
melt_norm_fus$norm_counts <- log2(melt_norm_fus$norm_counts)
melt_norm_fus$group <- factor(melt_norm_fus$group, levels = c('Normal','KD'))
melt_norm_fus$new_treatment <- factor(melt_norm_fus$new_treatment, levels = c('DMSO','abemaciclib','DMSO_RB1-KD','abemaciclib_RB1-KD'))

# Go to ggplot
```

# ggplot for all
```{r}
# Just change the parameter accordingly

ggplot(melt_norm, aes(x = new_treatment, y=norm_counts)) +
  geom_boxplot(aes(color=new_treatment)) +
  facet_wrap(~hgnc_symbol, scales = "free_y") +
  labs(title = "DESeq2 All+Fusion DMSO_RB1-KD vs Abemaciclib_RB1-KD", x="Treatment", y="Log2Normalized_Counts") +
  theme_bw() + # reset the setting theme, so needs to be in the first theme
  guides(col = guide_legend(title='Treatment')) +
  theme(plot.title = element_text(hjust=0.5), axis.text.x = element_text(angle = 45,hjust=1))
```

# ggplot for fusion
```{r}
# Adjust parameter accordingly

ggplot(melt_norm_fus, aes(x = new_treatment, y=norm_counts)) +
  geom_boxplot(aes(color=new_treatment)) +
  facet_wrap(~hgnc_symbol, scales = "free_y") +
  ggtitle("DESeq2 Fusion DMSO_RB1-KD vs Abemaciclib_RB1-KD") +
  theme_bw() + # reset the setting theme, so needs to be in the first theme
  guides(col = guide_legend(title='Treatment')) +
  theme(plot.title = element_text(hjust=0.5), axis.text.x = element_text(angle = 45,hjust=1)) +
  xlab("Treatment") +
  ylab("Log2Normalized_Counts")
```

# Volcano plot
```{r}
library(ggrepel)
ggplot(df.res, aes(x = log2FoldChange, y=-log10(pvalue), color = sig)) +
  geom_point() +
  xlim(-3.5,3.5) +
  ylim(0,40) +
  geom_text_repel(aes(label=hgnc.sig), size=2) +
  scale_color_manual(values=c('red','blue'), na.value = 'grey') +
  ggtitle("Volcano Plot DMSO vs DMSO_RB1-KO") +
  theme_bw() + # reset the setting theme, so needs to be in the first theme
  theme(plot.title = element_text(hjust=0.5)) +
  xlab("Log Fold Change") +
  ylab("-log10 p-Value")
```
